---
title: "Question One: Where are the volcanic eruptions likely to occur?"
output: html_notebook
---

# Import Libraries
```{r}
library(tidyverse)
library(leaflet)
library(ggplot2)
library(gapminder)
library(lubridate)
library(here)
library(sf)
library(rnaturalearth)

# library(rnaturalearthdata)
# library(rgeos)
# library(geojsonio)
# library(rgdal)
# library(ggmap)
# library(dplyr)
# library(hrbrthemes)
# library(viridis)
# library(RColorBrewer)
```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
```

```{r}
ggplot(data = world) +
  geom_sf(data = tectonic_plates,
          colour = "slateblue4") +
  geom_sf(color = "gray68", fill = "ivory") +
  ggtitle("Location of Volcanoes",
          subtitle = "(and the tectonic plates)")+
  geom_point(
     data = volcanoes,
     colour = "firebrick4",
     shape = 24,
     fill = "firebrick4",
     size =1, 
     aes(longitude, latitude)
      ) 
```
It is clear that volcanoes are located along the edges of the tectonic plates.

What other factors affect the location of active volcanoes?

```{r}
# NUmber of eruptions by year and elevation
eruptions_vei %>% 
ggplot(aes(x=start_year, y=elevation, colour=vei)) + 
    geom_point(size=1.5, alpha = 0.35) +
    scale_colour_manual(values =  c("0" = "khaki", "1" = "gold1", "2" = "darkgoldenrod2",
          "3" = "darkorange2", "4" = "firebrick1", "5" = "firebrick", "6" = "firebrick4",
          "7" = "black")) +
    labs(
      x = "Year of eruption",
      y = "Elevation (metres)",
      title = "Number of Eruptions by Year and Elevation",
      subtitle = "Using VEI scale"
      ) +
    theme(panel.background = element_rect(fill = "black"))
```


```{r}
# Number of eruptions by year and elevation
volcanoes %>%
  ggplot( aes(x=primary_volcano_type, y=elevation, fill=primary_volcano_type)) +
    geom_boxplot(varwidth = TRUE, alpha=0.4) +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="firebrick", size=0.4, alpha=0.6, shape = 24) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Elevation of volcanoes - grouped by primary volcano type") +
    labs(x = "Primary Volcano Type", y = "Elevation") +
    theme(plot.title = element_text(vjust = 0.5, hjust=0.3, size = 15),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.3, size = 10),
        axis.text.y = element_text(vjust = 0.5, hjust=0.3, size = 10))
```

What are those straight lines that are occurring? Perhaps some volcanoes are repeatedly erupting? Let's have a look.
```{r}
count_of_counts %>% 
ggplot(aes(x=e_count, y= no_volc)) + 
    geom_line(size=1, colour ="firebrick") +
    labs(
      x = "Number of times a volcano has erupted",
      y = "Count of Volcanoes",
      title = "How many times does a volcano erupt?"
      ) 
```
The majority of volcanoes erupt rarely (from what we know)
- 27% have only erupted once 
- 72% 10 or fewer eruptions
- 83% 20 or fewer

But there are some volcanoes that we know erupt a lot.
8 of the volcanoes have known to have erupted at least 100 times.

```{r}
ggplot(data = world) +
  geom_sf(data = tectonic_plates,
          colour = "slateblue4") +
  geom_sf(color = "gray68", fill = "ivory") +
  ggtitle("Location of Volcanoes which have knowingly erupted over 100 times") +
  geom_jitter(
     data = volcanoes_over_100,
     colour = "firebrick4",
     shape = 24,
     fill = "firebrick4",
     size =3, 
     aes(longitude, latitude)
      ) 
```
