---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
```

Which volcanoes are likely to cause the most damage if they blow?

```{r}
volcano_deaths %>% 
  ggplot(aes(x = year, y = deaths, colour=vei)) +
  geom_point(size=3.5, 
             alpha = 1, 
             shape = 18) +   
  scale_colour_manual(values =  c("0" = "khaki", "1" = "gold1", "2" = "darkgoldenrod2",
          "3" = "darkorange2", "4" = "firebrick1", "5" = "firebrick", "6" = "firebrick4",
          "7" = "black")) +
  labs(
      x = "Year of eruption",
      y = "Number of deaths",
      title = "Number of Deaths Caused by Volanoes",
      subtitle = "(Some estimates)"
      ) +
    theme(panel.background = element_rect(fill = "slategray2"))
```

```{r}
volcano_deaths %>% 
  filter(year>1820) %>% 
  ggplot(aes(x = year, y = deaths, colour=vei)) +
  geom_point(size=3.5, 
             alpha = 1, 
             shape = 18) +   
  scale_colour_manual(values =  c("0" = "khaki", "1" = "gold1", "2" = "darkgoldenrod2",
          "3" = "darkorange2", "4" = "firebrick1", "5" = "firebrick", "6" = "firebrick4",
          "7" = "black")) +
  labs(
      x = "Year of eruption",
      y = "Number of deaths",
      title = "Number of Deaths Caused by Volanoes",
      subtitle = "(after 1820)"
      ) +
    theme(panel.background = element_rect(fill = "slategray2"))
```

+--------------------------+--------------+-----------------------------+
| Volcano Name             | Country      | Population within 10km      |
+:=========================+:=============+:===========================:+
| Michoacan-Guanajuato     | Mexico       | 5783287                     |
+--------------------------+--------------+-----------------------------+
| Tatun Volcanic Group     | Taiwan       | 5084149                     |
+--------------------------+--------------+-----------------------------+
| Campi Flegrei            | Italy        | 2234109                     |
+--------------------------+--------------+-----------------------------+
| Ilopango                 | El Salvador  | 2049583                     |
+--------------------------+--------------+-----------------------------+
| Hainan Volcanic Field    | China        | 1731229                     |
+--------------------------+--------------+-----------------------------+
| San Pablo Volcanic Field | Philippines  | 1349742                     |
+--------------------------+--------------+-----------------------------+
| Ghegham Volcanic Ridge   | Armenia      | 1265153                     |
+--------------------------+--------------+-----------------------------+
| Dieng Volcanic Complex   | Indonesia    | 1092929                     |
+--------------------------+--------------+-----------------------------+
| Auckland Volcanic Field  | New Zealand  | 1049110                     |
+--------------------------+--------------+-----------------------------+
| Rahat, Harrat            | Saudi Arabia | 1010115                     |
+--------------------------+--------------+-----------------------------+

```{r}
ggplot() +
  geom_sf(data = tectonic_plates,
          colour = "slateblue4") +
  geom_sf(data = world, color = "gray87", fill = "ivory") +
  ggtitle("Size of the population within 100km of volcanoes",
          subtitle = "Top 10 highlighted") +
 theme(legend.position = "none") +
  geom_point(
    data = volcanoes,
    aes(x = longitude, 
        y = latitude, 
        colour = population_within_100_km,
        size = population_within_100_km),
    alpha = 0.3
  ) + 
  scale_color_distiller(palette = "Spectral") +
  geom_jitter(
     data = popn_1mn_plus,
     colour = "darkgreen",
     shape = 24,
     fill = "darkgreen",
     size = 2.5 , 
     aes(longitude, latitude)
      ) 
```

```{r}
ggplot(data = world) +
  geom_sf(data = tectonic_plates,
          colour = "slateblue4") +
  geom_sf(color = "gray68", fill = "ivory") +
  ggtitle("VEI of the Last Volcanic Eruptions") +
  geom_point(
    data = eruptions_vei,
    aes(x = longitude, y = latitude, 
        color = vei,
        alpha = 0.5)) +
  scale_colour_manual(values =  c("0" = "khaki", "1" = "gold1", "2" = "darkgoldenrod2",
          "3" = "darkorange2", "4" = "firebrick1", "5" = "firebrick", "6" = "firebrick4",
          "7" = "black")) +
  guides(alpha = "none")
```



